---
title: "project - cohort - sampleMatrixType - comment"
author: "Your Name"
date: "The Date"
output: 
  html_document:
    css: ../makeConfig.css
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(knitr)
library(htmlTable)
library(dplyr)

library(fusion)
    ## If you do not have fusion run below
# remotes::install_github("phenological/nmr-parser")
# remotes::install_github("jwist/fusion")
library(mva.plots)
    ## If you do not have mva.plot run below
# remotes::install_github("phenological/mva-plots")
library(nmr.spectra.processing)
    ## If you do not have nmr.spectra.processing run below
# remotes::install_github("phenological/nmr-spectra-processing")

```

# Load spectra daE

## Step 1: set the file.path

```{r}
# If you have your data stored in the OneDrive ANPC dataset folder then use below by replacing projectName and cohortName to appropriate names
pathToDaE<-file.path(Sys.getenv("DATASETS"),"projectName","cohortName","DataElements")
  # Or you can navigate yourself to the folder where you store daE file
pathToDaE<-"~/Dataset/ProjectName/cohortName/DataElements/"
```

## Step 2: load the daE

```{r}
### Lets say your Urine NMR spectra daE is called "Urine1DSpectra.daE" then
da<-local(get(load(file.path(pathToDaE,"Urine1DSpectra.daE"))))

```

## Step 3: extract information from the daE

[da\@varName](mailto:da@varName){.email}: column name of the da\@.Data
(ppm in this daE)

da\@.Data: spectra data

[da\@obsDescr](mailto:da@obsDescr){.email}: A list of specta information

-   info: You should find the dataPath, sampleID, sampleType and more

-   proc: (SF: spectra frequency. PHC0: zero-order phase. PHC1: 1st
    order phase. SR: Spectra Reference)

-   params: all the acqus information

-   test_tests_comment: QC report results

-   test_tests_value: QC report in details

-   test_infos_value: QC experiment performance results

```{r}

X<-da@.Data # spectra
ppm<-as.numeric(da@varName) 
Anno<-da@obsDescr$info

# if you put ppm to the column name of X, also sampleID to the row names would help you with plotting (viewing) later on
rownames(X)<-Anno$sampleID
colnames(X)<-ppm

```

# Post processing

## Step 1: calibrate spectra to alanine

### check the alanine spectra region (1.45 \~ 1.52 ppm) by plotting the regions using matspec()

<https://github.com/phenological/mva-plots/blob/main/R/matspec.R>

```{r}
matspec(X,ppm,roi = c(1.45,1.52),interactive = F)
```

### Calibrate to Alanine doublet using calibrateSpectra()

<https://github.com/phenological/nmr-spectra-processing/blob/main/R/calibrateSpectra.R>

```{r}
X_cal<-calibrateSpectra(ppm = ppm,Y = X,ref = c("alanine"),maxShift = 1.5,rOref = c(1.45,1.53),cshift = 1.48)
colnames(X_cal)<-ppm

matspec(X_cal,ppm,roi = c(1.45,1.52),interactive = F)

```

## Step 2: remove selected region of the spectra

```{r}

idx<-c(which(ppm<=0.5),which(ppm>=4.7 & ppm<=5.0),which(ppm>=9.5))

X<-X_cal[,-idx]
ppm<-ppm[-idx]

rm(idx,X_cal)
```

## Step 3: Apply baseline correction using baselineCorrection()

<https://github.com/phenological/nmr-spectra-processing/blob/main/R/baselineCorrection.R>

```{r}
X<-baselineCorrection(X)
```

## Step 4: Apply PQN normalization using normalize()

```{r}
X<-normalize(X)
```

### plot some spectra

Some region may looked aligned but in Urine, even after alignment to
alanine, you will find many miss aligned regions

```{r}
colnames(X)<-ppm
matspec(X,ppm,roi = c(5.2,5.3),interactive = F)
matspec(X,ppm,roi = c(7.0,7.5),interactive = F)

```

# Spectra check with PCA

You should see the "ltr" to be cluster tightly.

If not, look at the spectra and find out why (could it be failed water
suppression?)

You can also check if there are any outliers. Find out which spectra it
is, plot some spectra to find the reason why.

```{r}
mod<-PCA(X,rank = 2)
Anno$sampleType<-ifelse(grepl("COV",Anno$sampleID),"sample","ltr")

plotScores(mod,optns = list(color = Anno$sampleType))

PlotLoadSpec(mod,PC= 1,roi = c(0.5,4.5))
PlotLoadSpec(mod,PC= 1,roi = c(5.0,9.5))

```

# Remove LTRs, Outliers and Original run that has re-runs

If the LTRs are in tight cluster then we can remove them

Outliers: select those that should be excluded (e.g., experimental
errors)

Original runs of re-runs : when there is a re-runs it must be a reason..
you can check by plotting them to compare. In general, we want to keep
the last runs instead of the original runs.

You can find which one is reruns by looking at the sampleID. Those with
rerun should have "\#" or "." after the sampleID followed by the re-run
number. (e.g., COV00001#1)

Also noted that apart from Busselton project, all the sampleID should
have unique 3 capital letters given to each project at the beginning
followed by 5 digit numbers (total of 8 characters)

# Re-run Post Processing Step 1 to 4 without those LTR etc.,


# STOCSY
https://github.com/phenological/mva-plots/blob/main/R/stocsy.R

## Alanine (example)
d(1.48) 
q(3.78) signal intensity may be too low to see this compared to glucose peaks that are also overlap 
```{r}
d = 1.473# Driver peak in ppm
stocsy(x = ppm,Y = X,driver = d)
stocsy(x = ppm,Y = X,driver = d,roi = c(1.45,1.49))
# stocsy(x = ppm,Y = X,driver = d,roi = c(3.7,3.8))

```

## Glucose (example)
https://hmdb.ca/spectra/nmr_one_d/4817
```{r}
matspec(X[1:5,],ppm,roi = c(5.2,5.3))
d = 5.236# Driver peak in ppm
stocsy(x = ppm,Y = X,driver = d)
stocsy(x = ppm,Y = X,driver = d,roi = c(5.2,5.3))
stocsy(x = ppm,Y = X,driver = d,roi = c(4.6,4.7))
stocsy(x = ppm,Y = X,driver = d,roi = c(3.2,3.95))

```
